Automatic Parallelization and Transparent Fault Tolerance\\ (New project paper)

### 2016-04-05 pubs/tfp16/paper.tex
more small tweaks

Mutual exclusion for garbage collector--heap check with every heap allocation.
In fact garbage collector can be invoked at any point in time except when
populating a heap object immediately after its allocation.

@@

We perform standard Hindley-Milner type inference and
enforcement~\cite{Heren02}, augmented with built-in and user-defined unboxed
types as described by Peyton-Jones and Launchbury~\cite{Jone:1991},
but for unboxed types code generation and runtime support is currently limited
to built-in single-word (64-bit) or smaller types (e.g., \texttt{Int\#},
\texttt{UInt\#}, \texttt{Float\#}, \texttt{Double\#}) and user-defined
enumerations, e.g., \texttt{data unboxed Bool\# = False\# | True\#}.

@@

\subsubsection{\emph{Strict but still pure} is not new.}
Very simple implementations have been prototyped, for example
Sheard's.~\cite{Sheard:2003}
%
A plug-in for GHC (prior to 8.0) exists for making functions, but not data
structures, strict.~\cite{strict-ghc-plugin}.  The transformation is on GHC
Core, recursively expand non-recursive let bindings into case bindings.  Thus
it seems that while it makes functions denotationally strict, function arguments
are still evaluated after function call.
%
The Disciplined Disciple Compiler project appears to intend implementation of
a strict, higher order functional language (or family of languages) compiler
that has been underway for some time, but the emphasis appears to be on a much
more complex type system for handling stateful computation, and the prospects
for its eventual completion are unknown~\cite{disciplined-disciple}.
%
NVIDIA's NOVA language, and corresponding compiler, implements a strict
(``call by value''), pure, polymorphic, higher-order functional language with
built-in parallel operators (e.g., \emph{map}, \emph{reduce}, and
\emph{scan}).~\cite{NVIDIA:2013} Three back-ends were implemented: for
sequential C, parallel (multi-threaded) C, and CUDA C, the lattermost
targeting NVIDIA GPUs.

@@

Currently we are developing the parallel runtime.  This summer (2016) a
returning student will (partially) bridge the gap between GHC STG and our STG,
and augment our runtime as needed in this effort.  Just as in the Intel Haskell
Research Compiler effort, we will not

>>>>

Mutual exclusion for garbage collector--heap check with every heap allocation.
In fact the garbage collector can be invoked at any point in time except when
populating a heap object immediately after its allocation.

@@

We perform standard Hindley-Milner type inference and
enforcement~\cite{Heren02}, augmented with built-in and user-defined unboxed
types as described by Peyton-Jones and Launchbury~\cite{Jones:1991},
but for unboxed types code generation and runtime support is currently limited
to built-in single-word (64-bit) or smaller types (e.g., \texttt{Int\#},
\texttt{UInt\#}, \texttt{Float\#}, \texttt{Double\#}) and user-defined
enumerations, e.g., \texttt{data unboxed Bool\# = False\# | True\#}.

@@

\subsubsection{\emph{Strict but still pure} is not new.}
Very simple implementations have been prototyped, for example
Sheard's.~\cite{Sheard:2003}
%
A plug-in for GHC (prior to 8.0) exists for making functions, but not data
structures, strict.~\cite{strict-ghc-plugin}.  The transformation is on GHC
Core, recursively expanding non-recursive let bindings into case bindings.  Thus
it seems that while it makes functions denotationally strict, function arguments
are still evaluated after function calls.
%
The Disciplined Disciple Compiler project appears to intend implementation of
a strict, higher order functional language (or family of languages) compiler
that has been underway for some time, but the emphasis appears to be on a much
more complex type system for handling stateful computation, and the prospects
for its eventual completion are unknown~\cite{disciplined-disciple}.
%
NVIDIA's NOVA language, and corresponding compiler, implements a strict
(``call by value''), pure, polymorphic, higher-order functional language with
built-in parallel operators (e.g., \emph{map}, \emph{reduce}, and
\emph{scan}).~\cite{NVIDIA:2013} Three back-ends were implemented: for
sequential C, parallel (multi-threaded) C, and CUDA C, the lattermost
targeting NVIDIA GPUs.

@@

Currently we are developing the parallel runtime.  This summer (2016) a
returning student will (partially) bridge the gap between GHC STG and our STG,
and augment our runtime as needed in this effort.  Just as in the Intel Haskell
Research Compiler effort, we will use GHC as the front-end.
